---
title: 'DAND_P6: Data Visualization - Sketch V1'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Cyclical Trends of the US Housing Market (1963 - 2016)

An initial sketch from my exploratory data analysis showing the US Housing Boom/Bust Cycles.

__Narative: The US Housing market experiences periodic boom / bust cycles.  We are currently still recovering and a imminent bust is unlikely.__

___Proposed Development:___ 

1. Convert to dimple.js

2. The red and green vertical lines represent the Bust and Boom infliction points. Display the chart without these lines

3. Add an animation which adds each line sequentially.

4. Add 'Regional' buttons to view the data for US, NorthEast, MidWest, South, and West.

5. Add interactive time scaling / zooming

6. The "Sold" data as shown is smoothed with 12 month moving average because house sales are cyclical within each year.  However, at a certain zoom level, I will turn off the smoothing to allow the view to see these micro-trends.

7. Clean up the labels, scales and legends

__Any and all feedback is appreciated__



```{r Install Packages, include=FALSE}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This Dollarsprevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

#install.packages("scales")
#install.packages('ggplot2', dep = TRUE)
#install.packages("plyr")
#install.packages("tidyr")
#install.packages('alr3')
#install.packages("gridExtra")
#install.packages('reshape2')
#install.packages('GGally')
#install.packages('memisc')
#install.packages("readr")
#install.packages("dtplyr")
#install.packages("zoo")
#install.packages("smooth")


```

```{r, include=FALSE}
library(ggplot2)
library(tidyr)
library(alr3)
library(gridExtra)
library(GGally)
library(memisc)
library(reshape2)
library(readr)
library(memisc)
library(dtplyr)
library(zoo)
```

```{r, include=FALSE}
# Load the Data
setwd("C:/Users/rl1891/version-control/Visualizations/DAND_P6")
library(readr)
rs <- read_csv("C:/Users/rl1891/version-control/Visualizations/DAND_P6/rs.txt", 
    col_types = cols(cat_idx = col_skip(), 
        date = col_date(format = "%m/%d/%Y"), 
        dt_idx = col_skip(), geoID = col_factor(levels = c("US", 
            "NE", "MW", "SO", "WE")), geo_idx = col_skip(), 
        is_adj = col_skip(), is_adj1 = col_skip(), 
        prefix = col_skip(), suffix = col_skip(), 
        variable = col_factor(levels = c("AVERAG_SOLD", 
            "COMPED_FORSALE", "COMPED_SOLD", 
            "MEDIAN_SOLD", "MMTHS_FORSALE", 
            "MONSUP_FORSALE", "NOTSTD_FORSALE", 
            "NOTSTD_SOLD", "TOTAL_ASOLD", 
            "TOTAL_FORSALE", "TOTAL_SOLD", 
            "UNDERC_FORSALE", "UNDERC_SOLD"))))

```

```{r Import Dataset, include=FALSE}
# Convert Long Format to wide format
rsl <- dcast(rs, per_idx + date + geoID ~ variable, 
             fun.aggregate = mean, value.var = "val")
# rename the variables to more meaningful descriptions
names(rsl)[names(rsl)=="AVERAG_SOLD"] <- "mean_price_sold"
names(rsl)[names(rsl)=="COMPED_FORSALE"] <- "completed_houses_forsale"
names(rsl)[names(rsl)=="COMPED_SOLD"] <- "completed_houses_sold"
names(rsl)[names(rsl)=="MEDIAN_SOLD"] <- "median_price_sold"
names(rsl)[names(rsl)=="MMTHS_FORSALE"] <- "median_months_forsale"
names(rsl)[names(rsl)=="MONSUP_FORSALE"] <- "months_supply"
names(rsl)[names(rsl)=="TOTAL_ASOLD"] <- "annual_rate_sold"
names(rsl)[names(rsl)=="NOTSTD_FORSALE"] <- "not_started_forsale"
names(rsl)[names(rsl)=="NOTSTD_SOLD"] <- "not_started_sold"
names(rsl)[names(rsl)=="TOTAL_FORSALE"] <- "houses_forsale"
names(rsl)[names(rsl)=="TOTAL_SOLD"] <- "house_sold"
names(rsl)[names(rsl)=="UNDERC_FORSALE"] <- "under_construction_forsale"
names(rsl)[names(rsl)=="UNDERC_SOLD"] <- "under_construction_sold"

# Add year & decade to aid binning
rsl$year <- as.numeric(format(rsl$date,'%Y'))
rsl$decade <- round(as.numeric(format(rsl$date,'%Y'))/10)*10



```

```{r Make Plot, echo=FALSE, message=FALSE, warning=FALSE}
library(scales)
library(smooth)
library(zoo)
library(Mcomp)


# Create a subset of just US values
rsl_us <- subset(rsl, rsl$geoID == "US")

new <- data.frame(date=rsl_us$date, 
                  forsale_pct=rsl_us$houses_forsale / mean(rsl_us$houses_forsale), 
                  sold_pct = ave(rsl_us$house_sold, 
                                 FUN= function(x) rollmean(x, k=12, 
                                  fill = NA ))/ mean(as.double(rsl_us$house_sold)))



rsl3 <- melt(new[,c("date","forsale_pct", "sold_pct")], 
                id.vars = "date")

cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
                "#0072B2", "#D55E00", "#CC79A7")

a = 1.0
color1 = "green"
color2 = "red"

# Create a list of vertical lines for the critcal dates to be added to plots
p_critcal = list(geom_vline(xintercept = 
                            as.numeric(as.Date('1/1/1976',format='%m/%d/%Y')),
                            color = color1, alpha = a),
                 geom_vline(xintercept = 
                            as.numeric(as.Date('6/1/1979',format='%m/%d/%Y')),
                            color = color2, alpha = a), 
                 geom_vline(xintercept = 
                            as.numeric(as.Date('6/1/1982',format='%m/%d/%Y')),
                            color = color1, alpha = a),
                 geom_vline(xintercept = 
                            as.numeric(as.Date('6/1/1989',format='%m/%d/%Y')),
                            color = color2, alpha = a),
                 geom_vline(xintercept = 
                            as.numeric(as.Date('10/1/1992',format='%m/%d/%Y')),
                            color = color1, alpha = a),
                 geom_vline(xintercept = 
                            as.numeric(as.Date('1/1/1996',format='%m/%d/%Y')),
                            color = color2, alpha = a),
                 geom_vline(xintercept = 
                            as.numeric(as.Date('1/1/1998',format='%m/%d/%Y')),
                            color = color1, alpha = a),
                 geom_vline(xintercept = 
                            as.numeric(as.Date('6/1/2006',format='%m/%d/%Y')),
                            color = color2, alpha = a),
                 geom_vline(xintercept = 
                            as.numeric(as.Date('4/1/2012',format='%m/%d/%Y')),
                            color = color1, alpha = a))

plt <- ggplot(rsl3, aes(date, value, color = variable)) + 
  geom_line(size = 1) + 
  geom_hline(aes(yintercept=1.0), color="black", linetype="dashed", size=.5) +
  ylab('Houses: For Sale & Sold Percent of Mean') +
  labs(color="Legend:") +
  ggtitle("House ForSale, Sold Percent of Mean by Date (1963-2016)") +
  scale_x_date(date_breaks = "5 year",
               labels=date_format("%Y"),
               limits = as.Date(c('1973-01-01','2016-12-31')))
plt + p_critcal + theme_bw()

#ggsave("DAND_P6_sketch_v1.pdf")

```

__Dataset Information:__

- url: http://wwww.census.gov/econ/currendata/datasets

- "Economic Measures Time Series Dataset"




